---
title: "extract fetch pixels"
author: "Gavin McNicol"
date: "2026-01-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 8,
  fig.height = 5
)

# Prefer an explicit project root (assumes you knit from the Review/ folder or project root)
# If you use an RStudio Project, set your working directory to the project root before knitting.
```

## Purpose: Review (format, content, visualization) the final output of the WetLSP processing pipeline.

```{r libraries}
required_packages <- c(
  "terra", "sf", "tidyverse", "rjson", "geojsonR", "ncdf4"
)

load_or_stop <- function(pkgs) {
  missing <- pkgs[!vapply(pkgs, requireNamespace, FUN.VALUE = logical(1), quietly = TRUE)]
  if (length(missing) > 0) {
    stop(
      "Missing packages: ", paste(missing, collapse = ", "),
      "\nInstall them (once) via install.packages(c(...)) and re-knit."
    )
  }
  invisible(lapply(pkgs, library, character.only = TRUE))
}

load_or_stop(required_packages)

source("../pipeline/validation/wetlsp-helpers.R")
```


## REVIEW

## 1. Load and basic structure checks

NOTE: for multi-part RDS sites use this chunk first

```{r}
site_name <- "GF-Guy"
site_dir <- file.path("../output", site_name)

part_paths <- list.files(
  path   = file.path(site_dir, out_subdir),
  pattern = paste0(
    "^", site_name,
    "-evi-timeseries-fetch-sf-",
    radius_m, "m-part\\d+of\\d+\\.rds$"
  ),
  full.names = TRUE
)

stopifnot(length(part_paths) > 1)
part_paths <- sort(part_paths)
part_paths
```

## or for one rds

```{r}
site_name <- "FR-LGt"
radius_m <- 200
site_dir <- file.path("../output", site_name)
out_subdir <- paste0("tables_sf_fetch_", radius_m, "m")

sf_all_path <- file.path(site_dir, out_subdir, paste0(site_name, "-evi-timeseries-fetch-sf-", radius_m, "m.rds"))

sf_all <- readRDS(sf_all_path)

sf_all
st_crs(sf_all)
```

## 2. Spatial footprint + fetch circle overlay

```{r}
library(ggplot2)

ggplot(sf_all) +
  geom_sf(size = 0.3, alpha = 0.6) +
  coord_sf() +
  labs(
    title = paste0(site_name, " fetch-limited EVI pixels"),
    subtitle = paste("n =", nrow(sf_all))
  ) +
  theme_minimal()
```



## 3. Pixel density & geometry sanity

```{r}
# Are there duplicate geometries?
any(duplicated(st_as_binary(st_geometry(sf_all))))

# Bounding box size (meters)
st_bbox(sf_all)
```

## 4. Time-series presence checks (very important)

```{r}
summary(lengths(sf_all$evi_raw))
summary(lengths(sf_all$evi_spline))

mean(lengths(sf_all$evi_raw) > 0)
mean(lengths(sf_all$evi_spline) > 0)
```

## 5. Pick a few random pixels and plot their time series

```{r}
PickPixelInteractive(sf_all)
ViewPixelTimeseries(sf_all, cell_vec = c(seq(676446, 676846, by = 5)))
ViewPixelViolin(sf_all, cell_vec = c(675123, 617804, 677831, 671202))
ViewPixelBoxplot(sf_all, cell_vec = c(675123, 617804, 677831, 671202))

ViewPixelBoxplot(sf_all, cell_vec = c(725562))
ViewPixelTimeseries(sf_all, cell_vec = c(675123, 617804, 677831, 671202))
```

```{r}
library(tidyverse)
library(purrr)

## random pixels
# set.seed(1)
# pix <- sample(seq_len(nrow(sf_all)), 8)

# selected pixels
cell_vec <- c(
  "1 - Road"  = 619033,
  "2 - Clearing" = 965327,
  "3 - More Trees" = 82571,
  "4 - Trees"    = 565567
)

pix <- match(cell_vec, sf_all$cell)

plot_df <- purrr::map_dfr(seq_along(pix), function(j) {

  i <- pix[[j]]
  label <- names(cell_vec)[[j]]
  cell_id <- cell_vec[[j]]

  tibble(
    pixel = label,   # <- human-readable name
    cell  = cell_id, # optional, keep for reference
    date  = c(
      as.Date(sf_all$dates_spline[[i]]),
      as.Date(sf_all$dates_raw[[i]])
    ),
    evi   = c(
      sf_all$evi_spline[[i]],
      sf_all$evi_raw[[i]]
    ),
    type  = c(
      rep("Spline",   length(sf_all$evi_spline[[i]])),
      rep("Observed", length(sf_all$evi_raw[[i]]))
    )
  )
})

# ---- plot ----
ggplot(plot_df, aes(x = date, y = evi)) +
  geom_line(
    data = filter(plot_df, type == "Spline"),
    color = "darkgreen",
    linewidth = 1.3
  ) +
  geom_point(
    data = filter(plot_df, type == "Observed"),
    color = "grey40",
    size = 1.6,
    alpha = 0.6
  ) +
  facet_wrap(~ pixel, ncol = 2, scales = "free_y") +
  labs(
    x = "Date",
    y = "EVI"
  ) +
  theme_bw() +
  scale_x_date(limits = c(as.Date(ymd("2021-01-01")), as.Date(ymd("2024-12-31")))) +
  scale_y_continuous(limits = c(0, 1)) +
  theme(
    strip.background = element_rect(fill = "grey95"),
    panel.grid.minor = element_blank()
  )
```






